Manuscript Figures

### 2024-06-24 code/Manuscript_figures.rmd
modify summary figures, try to generate sankey diagram but unsucessful

Summary plots
# Load required packages
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
# devtools::install_github("pedropark99/ggfunnel")
library(ggfunnel)
library(ggpubr)
```

```{r}
Step <- c("Raw", "After cleaning", "Full trait coverage")
frug_records <- c(8089513, 1721825, 1667892)
frug_species <- c(1222, 1219, 1074)
frug_genera <- c(400, 398, 370)
frug_families <- c(66, 66, 64)

frug_records_df <- data.frame(Step, frug_records)
frug_species_df <- data.frame(Step, frug_species)
frug_genera_df <- data.frame(Step, frug_genera)
frug_families_df <- data.frame(Step, frug_families)
```

```{r}
ggplot(frug_records_df, aes(x = Step, y = frug_records)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = frug_records), color = "white") + 
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(),axis.ticks.x = element_blank(), axis.title.y = element_blank())
```
```{r}
frug_records_fig <- ggfunnel::funnel(frug_records_df, levels = Step, values = frug_records, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore records", y = NULL)
frug_records_fig
```

```{r}
frug_species_fig <- ggfunnel::funnel(frug_species_df, levels = Step, values = frug_species, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore species", y = NULL)
frug_species_fig
```

```{r}
frug_genera_fig <- ggfunnel::funnel(frug_genera_df, levels = Step, values = frug_genera, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore genera", y = NULL)
frug_genera_fig
```

```{r}
frug_families_fig <- ggfunnel::funnel(frug_families_df, levels = Step, values = frug_families, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore families", y = NULL)
frug_families_fig
```

```{r}
figure <- ggarrange(frug_records_fig,frug_species_fig, frug_genera_fig, frug_families_fig,
                    labels = c("A", "B", "C", "D"),
                    ncol = 2, nrow = 2)
figure
```

```{r}
devtools::session_info()
```

>>>>

---
title: "Manuscript Figures"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script generates misc figures for manuscript that corresponds to this project."
date: "2024-05-14"
output: 
    html_document:
      code_link: TRUE
---

# Set file paths
```{r}
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

# Load required packages
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggpubr)
library(networkD3)
#devtools::install_github("davidsjoberg/ggsankey")
library(ggsankey)
```

# Creating cleaning occurrence data figures

These numbers are recorded from the summary function at the end of most L0 and L1 scripts. A table is available in the manuscript supplemental materials.

## Frugivores
```{r}
Step <- c("Raw", "After cleaning", "Full trait coverage")
frug_records <- c(8885330, 1721825, 1667892)
frug_species <- c(1360, 1219, 1074)
frug_genera <- c(400, 398, 370)
frug_families <- c(66, 66, 64)

frug_summary_df <- data.frame(Step, frug_records, frug_species, frug_genera, frug_families)

# Reshape data to long format suitable for plotting
frug_summary_long <- frug_summary_df %>%
  pivot_longer(cols = starts_with("frug"), names_to = "Category", values_to = "Value")
```

```{r}
# Define the order of steps with "Raw" at the top
step_order <- c("Full trait coverage", "After cleaning", "Raw")

# Convert Step to factor with defined order
frug_summary_long$Step <- factor(frug_summary_long$Step, levels = step_order)

# Replace Category labels with more descriptive names
frug_summary_long <- frug_summary_long %>%
  mutate(Category = case_when(
    Category == "frug_records"   ~ "Records",
    Category == "frug_species"   ~ "Species",
    Category == "frug_genera"    ~ "Genera",
    Category == "frug_families"  ~ "Families",
    TRUE                         ~ as.character(Category)  # Handle any other cases
  ))

# Define the order of categories
category_order <- c("Records", "Species", "Genera", "Families")

# Convert Category to factor with defined order
frug_summary_long$Category <- factor(frug_summary_long$Category, levels = category_order)

```

```{r}
(frugivore_summary_plot <- ggplot(frug_summary_long, aes(y = Step, x = Value, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(. ~ Category, scales = "free_x") +
  labs(title = "Frugivore Data Summary",
       x = "Value",
       y = NULL) +
  theme_minimal() +
  theme(strip.background = element_blank(),  # Remove background from facet labels
        strip.text = element_text(size = 12, face = "bold"),
        legend.position = "none"))
ggsave("frugivore_summary_plot.png", frugivore_summary_plot, path = figure_path)
```

## Plants

```{r}
Step <- c("Raw", "After cleaning", "Full trait coverage")
plant_records <- c(4519792, 1886120, 1700044)
plant_species <- c(40288, 39972, 33746)
plant_genera <- c(4083, 3793, 3400)
plant_families <- c(307, 293, 275)

plant_summary_df <- data.frame(Step, plant_records, plant_species, plant_genera, plant_families)

# Reshape data to long format suitable for plotting
plant_summary_long <- plant_summary_df %>%
  pivot_longer(cols = starts_with("plant"), names_to = "Category", values_to = "Value")
```

```{r}
# Define the order of steps with "Raw" at the top
step_order <- c("Full trait coverage", "After cleaning", "Raw")

# Convert Step to factor with defined order
plant_summary_long$Step <- factor(plant_summary_long$Step, levels = step_order)

# Replace Category labels with more descriptive names
plant_summary_long <- frug_summary_long %>%
  mutate(Category = case_when(
    Category == "plant_records"   ~ "Records",
    Category == "plant_species"   ~ "Species",
    Category == "plant_genera"    ~ "Genera",
    Category == "plant_families"  ~ "Families",
    TRUE                         ~ as.character(Category)  # Handle any other cases
  ))

# Define the order of categories
category_order <- c("Records", "Species", "Genera", "Families")

# Convert Category to factor with defined order
plant_summary_long$Category <- factor(plant_summary_long$Category, levels = category_order)

```

```{r}
(plant_summary_plot <- ggplot(plant_summary_long, aes(y = Step, x = Value, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(. ~ Category, scales = "free_x") +
  labs(title = "Plant Data Summary",
       x = "Value",
       y = NULL) +
  theme_minimal() +
  theme(strip.background = element_blank(),  # Remove background from facet labels
        strip.text = element_text(size = 12, face = "bold"),
        legend.position = "none"))
ggsave("plant_summary_plot.png", plant_summary_plot, path = figure_path)
```

# Create Sankey diagram for taxonomic harmonization
Inspiration figures: https://conbio-onlinelibrary-wiley-com.proxy2.cl.msu.edu/doi/10.1111/csp2.311#csp2311-fig-0003 and https://bsapubs-onlinelibrary-wiley-com.proxy2.cl.msu.edu/doi/full/10.1002/aps3.11575#aps311575-fig-0003

Package used (databases)    Plant species inputted    Matches   Disparities
TNRS                        487,201                   472,680   14,548
bdc (COL)                   14,548                    5,177     9,371

*The following code doesn't generate the figure like I want.*
```{r}
# Create connection dataframe
links <- data.frame(
  source = c("Plant species inputted", "TNRS", "TNRS", "Disparities", "bdc (COL)", "bdc (COL)", "Matches"),
  target = c("TNRS", "Matches", "Disparities", "bdc (COL)", "Matches", "Species dropped", "Species harmonized"),
  value = c(487201, 472680, 14548, 14548, 5177, 9371, 476010)
)

# Create nodes
nodes <- data.frame(
  name = c(as.character(links$source),
           as.character(links$target) %>% unique())
)

# Reformat connections
links$IDsource <- match(links$source, nodes$name)-1
links$IDtarget <- match(links$target, nodes$name)-1

# Create the Sankey diagram
sankeyNetwork(Links = links, Nodes = nodes, Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name")
```

```{r}
# Create the data frames for the nodes and links
sankey_data <- data.frame(
  x = c("source", "target", "target", "source", "target", "target", "source", "target", "source"),
  node = c("Plant species inputted", "TNRS", "TNRS", "Disparities", "bdc (COL)", "bdc (COL)", "Species dropped", "Matches", "Species harmonized"),
  next_x = c("target", "source", "source", "target", "source", "source", NA, "source", NA),
  next_node = c("TNRS", "Matches", "Disparities", "bdc (COL)", "Matches", "Species dropped", NA, "Species harmonized", NA),
  value = c(487201, 472680, 14548, 14548, 5177, 9371, 9371, 476010, 476010)
)

```

```{r}
# Create the Sankey diagram
ggplot(sankey_data, aes(x = x, next_x = next_x, node = node, next_node = next_node, value = value)) +
  geom_sankey(flow.fill = "darkgrey", node.fill = "darkblue", node.color = "darkblue", alpha = 0.5) +
  geom_sankey_label(aes(label = node), size = 3, fill = "white") +
  theme_sankey(base_size = 12) +
  theme_void() +
  labs(title = "Plant Species Matching Workflow",
       subtitle = "Visualization of species input, TNRS and BDC matching process",
       x = NULL,
       y = NULL) +
  theme(legend.position = "none")
```

# References

```{r}
library(report)
cite_packages()
```

```{r, class.source= 'fold-hide'}
devtools::session_info()
```
